---
title: "Births 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This dataset collected information on United States births in 2018. This dataset can be downloaded for **free** using Kaggle.com. 

##### Link 
https://www.kaggle.com/des137/us-births-2018/version/2

##### User Guide
https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/DVS/natality/UserGuide2018-508.pdf

## Import Dataset

```{r, message=FALSE}
library(readr)
us.births <- read_csv("~/Documents/DA/US_births(2018).csv")
```

## Create Dateframe Containing Targeted Variables 

```{r}
df.births <- data.frame(us.births$MAGER, us.births$MRACE15, us.births$DMAR, us.births$MEDUC, us.births$PREVIS, us.births$RF_CESAR, us.births$RF_CESARN, us.births$LD_INDL, us.births$RDMETH_REC, us.births$ATTEND, us.births$PAY_REC)
```

### Targeted Variables

* MAGER
     + Age of Birthing Person
* MRACE15
     + Race of Birthing Person
* DMAR
     + Marital Status of Birthing Person
* MEDUC
     + Education Description of Birthing Person
* PAY_REC
     + Payment Method 
* PREVIS
     + Number of Prenatal Visits
* RDMETH_REC
     + Method of Delivery
* RF_CESAR
     + Previous Cesarean Section Recorded
* RF_CESARN
     + Number of Previous Cesarean Section Births Recorded
* ATTEND
     + Birth Attendant During Delivery
* LD_INDL
     + Labor Induction Method Recorded

## Cleaning Dateframe

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
str(df.births)
summary(df.births)
```

```{r, mager}
df.births$us.births.MAGER<- as.numeric(df.births$us.births.MAGER)
```

```{r, mrace15}
df.births$us.births.MRACE15<- recode_factor(df.births$us.births.MRACE15,`1` = "White (only)", `2` = "Black (only)", `3` = "AIAN (only)", '4' = "Asian Indian (only)", '5' = "Chinese (only)", '6' = "Filipino (only)", '7' = "Japanese (only)", '8' = "Korean (only)", '9' = "Vietnamese (only)", '10' = "Other Asian (only)", '11' = "Hawaiian (only)", '12' = "Guamanian (only)", '13' = "Samoan (only)", '14' = "Other Pacific Islander (only)", '15' = "More than one race")
```

```{r, dmar}
df.births$us.births.DMAR<- recode_factor(df.births$us.births.DMAR, '1' = "Married", '2' = "Not Married")
```

```{r, meduc}
df.births$us.births.MEDUC[df.births$us.births.MEDUC==9]<- NA
df.births$us.births.MEDUC<- recode_factor(df.births$us.births.MEDUC, '1' = "8th grade or less", '2' = "9th through 12th grade with no diploma", '3' = "High school graduate or GED completed", '4' = "Some college credit, but not a degree.", '5' = "Associate degree (AA,AS)", '6' = "Bachelor’s degree (BA, AB, BS)", '7' = "Master’s degree (MA, MS, MEng, MEd, MSW, MBA)", '8' = "Doctorate (PhD, EdD) or Professional Degree (MD, DDS,
DVM, LLB, JD)")
```

```{r, payrec}
df.births$us.births.PAY_REC[df.births$us.births.PAY_REC==9]<- NA
df.births$us.births.PAY_REC<- recode_factor(df.births$us.births.PAY_REC, '1' =
"Medicaid", '2' = "Private Insurance", '3' = "Self Pay", '4' = "Other")
```

```{r, previs}
df.births$us.births.PREVIS[df.births$us.births.PREVIS==99]<- NA
df.births$us.births.PREVIS<- as.numeric(df.births$us.births.PREVIS)
```

```{r, methrec}
df.births$us.births.RDMETH_REC[df.births$us.births.RDMETH_REC==9]<- NA
df.births$us.births.RDMETH_REC<- recode_factor(df.births$us.births.RDMETH_REC,'1'= "Vaginal (excludes vaginal after previous C-section)", '2' = "Vaginal after previous c-section",'3' = "Primary C-section",'4' = "Repeat C-section", '5' = "Vaginal (unknown if previous c-section)", '6' = "C-section (unknown if previous c-section)")
```

```{r, csec}
df.births$us.births.RF_CESAR<- recode_factor(df.births$us.births.RF_CESAR, 'Y' = "YES", 'N' = "No", 'U' = "Unknown or not stated")
```

```{r, numcsec}
df.births$us.births.RF_CESARN<- as.numeric(df.births$us.births.RF_CESARN)
df.births$us.births.RF_CESARN[df.births$us.births.RF_CESARN==99]<- NA
```

```{r, attenrec}
df.births$us.births.ATTEND[df.births$us.births.ATTEND==9]<- NA
df.births$us.births.ATTEND<- recode_factor(df.births$us.births.ATTEND, '1' = "Doctor of Medicine (MD)", '2' = "Doctor of Osteopathy (DO)", '3' = "Certified Nurse Midwife (CNM)", '4' = "Other Midwife", '5' = "Other", '9' = "Unknown or not stated")
```

```{r, indrec}
df.births$us.births.LD_INDL<- recode_factor(df.births$us.births.LD_INDL, 'Y' = "YES", 'N' = "No", 'U' = "Unknown or not stated")
```

```{r}
str(df.births)
summary(df.births)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
